Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2023-01-11 edubots_surveys.Rnw
More revisions with the help of ChatGPT

Furthermore, the answers to the survey question regarding COVID highlights the
significant challenge of altering behavior, even during a major crisis. Very few
responders actually changed their habits, although comparing women vs. men as
well as those teaching social sciences vs. the rest of the disciplines, there
are some significant differences. We do not have any kind of explanation for
this behavior, other than technologies are perceived, and adopted, in a
different way depending on background and, apparently, gender.

One of the most interesting questions arising from the analysis of the results,
and that should require further study, is: Are teachers really aware of the
possibilities of chatbots in the classroom? The survey only listed a few
functionalities for the chatbots, and just a few subjects selected the option
``Others''. This may suggest that most teachers did not really understand at
this point the actual potential of chatbots, since the responses listed were by
no means exhaustive. If the responders were more acquainted with the
possibilities of this technology, this response would possibly have shown up
many more times. Another way to look at it is that, unless you provide the
specific instructions on how to use a tool, the teaching staff is already too
busy or burnt out to think about innovation.

We can speculate that teaching staff's opinions of chatbots are influenced by
negative experiences with chatbots used in customer support, which is what most
people are likely to have experienced. Further study may be needed to confirm
this, but it is important to keep in mind when introducing chatbot technologies
in the classroom.  To potentially change this existing perception, it will be
necessary to make chatbots as simple and satisfying to use as possible.

Regarding gender grouping, both genders agree in their uses, so it seems not
relevant at all. However, when we look at the plot containing the sector
differentiation, it is interesting to note that university teaching staff are
more likely to use chatbots for grades and FAQs compared to other tertiary
education teaching staff.  This seems reasonable as the frequency of seeing the
students is lower at the University so, an automatic tool for grade notification
for continuous evaluation would be more helpful.  Furthermore, the high
student-to-teacher ratios that university educators often face could be another
reason why a higher number of them are willing to use chatbots for FAQs, as this
could reduce misunderstandings and email overload from students.

@@

According to the responses presented above, teachers’ responses are evenly split
between positive and negative in relation to the media features: analytics,
connectivity, and hidden phone number. This means that none of these features is
a deciding factor in their choice of technology. However, for features such as
familiarity, horizontality, sustainability, and officiality the majority
provided a positive response, designating their preference to these features. On
the other hand, the majority of teachers do not seem to see the media features
of official certification, pluggability, unidirectionality, and synchrony as
valuable in messaging apps.

From a qualitative point of view, these results show consensus in a few key aspects:
\begin{itemize}
\item Teaching staff want to have support from their institutions. This is
  probably due to the strict European data regulations on data protection, but
  also to avoid the overhead required to sign up new students every year.  In
  general, additional support with specific formation and access to IT help desk
  for specific features enhance the possibility of a successful adoption that
  improves the learning outcomes. This is supported by the result of the two
  surveys: they prefer whatever ``messaging'' application provided by as well as
  the fact that this application is the ``official'' for the university, as
  shown in Table \ref{tab:interaction}.
\item Teachers need sustainability. The main reason is that implementing changes
  is expensive in terms of effort and prone to errors in the first
  years. Therefore, it is logical that if chatbots or messaging platforms are to
  be introduced, teaching staff want this change to be as permanent as possible.
 \item Keep it Simple. Media interaction requirements are simple and only 2.4\%
   of educators are requesting more features. This could be a sign of
   technological burnt out produced by the previous pandemic courses, when the
   use of computers significantly increased.
\end{itemize}

The analysis of the questions related to RQ4 seems to indicate that teaching
staff prefer to share an interaction space with their students. As shown in
Tables \ref{tab:kindinteractions} and \ref{tab:interaction}, the majority of
responders indicated that they {\em do not} want unidirectional communication,
and that they prefer {\em horizontal} interaction. This point is reinforced by
results on the kind of chatbots they preferred: most of them indicated FAQ bots,
which might indicate a need to offload part of the burden of answering every
single question posed by their students. In a horizontal setting, other students
would answer if the teacher does not do it immediately. This might also be the
reason why the possibility of using tools from which analytics can be extracted
is also valued by teachers: in that way, the student activity in the chat tool,
answering questions, or their regularity can all be assessed as part of the
process of achieving learning objectives.

In our survey, we could not find an answer about the right moment to introducing
chatbot technologies. The response to questions regarding to change of behavior
during and after the first stay-at-home stage of the pandemic do not suggest
that a major (or minor, such as new higher education laws\footnote{We seem to
  have one of these, at least in Spain, every 10 years or so.} or the
introduction of new degrees) crisis could be an opportunity to introduce new
technologies. The reason being that it does not bring major changes in attitude
Although changes driven by external forces do offer the chance of piloting new
technologies, they do not seem to bring internal changes in educators' attitudes
(which, after all, is bound to be the same). In absence of a clear answer in
this direction, the right moment to introduce new technologies is always when
the IT and managing staff is ready to support it (since, as we have seen before,
``official'' support is one of the factors that is most valued by teachers).

>>>>

Furthermore, the answers to the survey question regarding COVID highlights the
significant challenge of altering behavior, even during a major crisis. Very few
responders actually changed their habits, although comparing women vs. men as
well as those teaching social sciences vs. the rest of the disciplines, there
are some significant differences. We do not have a clear understanding of the
reasons behind this behavior; however, it appears that the perception and
adoption of technologies in a major crisis may vary based on an individual's
background and gender.

One of the most interesting questions arising from the analysis of the results,
and that should require further study, is: Are teachers really aware of the
possibilities of chatbots in the classroom? The survey only listed a few
functionalities for the chatbots, and just a few subjects selected the option
``Others''. This may suggest that most teachers did not really understand at
this point the actual potential of chatbots, since the responses listed were by
no means exhaustive. Had the responders been more acquainted with the
possibilities of this technology, this ``Other'' response would possibly have
shown up many more times. Another way to look at it is that, unless you provide
the specific instructions on how to use a tool, the teaching staff is already
too busy or burnt out to think about innovation, so they think it is not their
job to think about new possibilities of existing or new technologies.

The reluctance to adopt chatbots in class might have an explanation: We can
speculate that teaching staff's opinions of chatbots are influenced by negative
experiences with chatbots used in customer support, which is what most people
are likely to have experienced. Further study may be needed to confirm this, but
it is important to keep it in mind when introducing chatbot technologies in the
classroom.  To potentially change this existing perception, it will be necessary
to make chatbots as simple and satisfying to use as possible.

Regarding gender grouping, both genders agree in the possible uses of chatbots,
so it seems not relevant at all. However, when we look at the plot containing
the sector differentiation, it is interesting to note that university teaching
staff are more likely to use chatbots for grades and FAQs compared to teaching
staff involved in other tertiary education institutions. This appears logical,
as the number of personal interactions with students in certain classes at the
university is smaller\footnote{Most university classes take place one or twice
  week, as opposed to daily classes with the same teacher in other tertiary
  education, and hold usually twice as many persons as the typical tertiary
  education class}, thus an automated grade notification tool for continuous
evaluation would be more beneficial.  Furthermore, the high student-to-teacher
ratios that university educators often face could be another reason why a higher
number of them are willing to use chatbots for FAQs, as this could reduce
misunderstandings and email (or face to face tutoring) overload from students.

@@

According to the responses presented above, teachers’ reactions are evenly split
between positive and negative in relation to the media features: analytics,
connectivity, and hidden phone number. This means that none of these features is
a deciding factor in their choice of technology. However, for features such as
familiarity, horizontality, sustainability, and officiality the majority
provided a positive response, designating their preference to these features. On
the other hand, the majority of teachers do not seem to see the media features
of official certification, pluggability, unidirectionality, and synchrony as
valuable in messaging apps.

From a qualitative point of view, these results show consensus in a few key aspects:
\begin{itemize}
\item Teaching staff wants to have support from their institutions. One of the
  main reasons why this is a factor might be the strict European data
  regulations on data protection; it is very likely that they also want to avoid
  the overhead required to sign up new students every year and other maintenance
  tasks.  In general, additional support with specific formation and access to
  IT help desk for specific features enhance the possibility of a successful
  adoption that improves the learning outcomes. This is supported by the result
  of the two surveys: they prefer whatever ``messaging'' application provided by
  as well as the fact that this application is the ``official'' for the
  university, as shown in Table \ref{tab:interaction}.
\item Teachers require sustainability. The primary rationale is that the
  implementation of changes is costly in terms of effort and prone to mistakes
  during the initial years of deployment of any new technology. Therefore, it is
  logical that if chatbots or messaging platforms are to be introduced, teaching
  staff want this change to be as permanent, and long-term, as possible.
 \item ``Keep it Simple''. Media interaction requirements are simple and only 2.4\%
   of educators are requesting more features. This could be a sign of
   technological burnout produced by the previous pandemic courses, when the
   use of computers significantly increased; however, simplicity is always a
   feature that people want to have in any technology they adopt.
\end{itemize}

The analysis of the answers to questions related to RQ4 seems to indicate that
teaching staff prefer to share an interaction space with their students. As
shown in Tables \ref{tab:kindinteractions} and \ref{tab:interaction}, the
majority of responders indicated that they {\em do not} want unidirectional
communication, and that they prefer {\em horizontal} interaction. This point is
reinforced by results on the kind of chatbots they preferred: most of them
indicated FAQ bots, which might indicate a need to offload part of the burden of
answering every single question posed by their students. In a horizontal
setting, other students would answer if the teacher does not do it
immediately. This might also be the reason why the possibility of using tools
from which analytics can be extracted is also valued by teachers: in that way,
the student activity in the chat tool, answering questions, or their regularity
can all be assessed as part of the process of achieving learning objectives.

In our survey, we could not find an answer about the right moment for the
introduction of chatbot technologies. The response to questions regarding to
change of behavior during and after the first stay-at-home stage of the pandemic
do not suggest that a major (or minor, such as new higher education
laws\footnote{We seem to have one of these, at least in Spain, every 10 years or
  so.} or the introduction of new degrees) crisis could be an opportunity to
introduce new technologies. The reason being that it does not bring major
changes in attitude. Although changes driven by external forces do offer the
chance of piloting new technologies, they do not seem to bring internal changes
in educators' attitudes (which, after all, is bound to be the same). In absence
of a clear answer in this direction, the right moment to introduce new
technologies is always when the IT and managing staff is ready to support it
(since, as we have seen before, ``official'' support is one of the factors that
is most valued by teachers).
